### 2016-03-10 thesis/conclusions.tex
all chapters but concl

Chapter \ref{ch:galviv} \citep{paperI} was dedicated to the careful multicomponent decomposition of 66 galaxies 
with a direct measurement of the black hole mass. 
We followed the same approach as \citet{laurikainen2005} in selecting the components for each galaxy model.  
\emph{A priori} identification of the galaxy components was done on the basis of several different indicators 
such as the analysis of the isophotal parameters, 
the inspection of unsharp masks, 
complementary information extracted from the literature, 
and -- most importantly -- from the galaxy kinematics. 
A joint photometric-kinematic approach (e.g.~\citealt{krajnovic2013,arnold2014}) 
turned out to play a decisive role for the robustness of our galaxy modelling. 
Upon examining central \citep{atlas3dIII,scott2014} and more extended \citep{arnold2014} velocity maps 
and comparing kinematic and photometric signatures of stellar discs, 
we were able to securely identify the presence and the radial extent of such discs. 
We observed a wide range of spheroid-to-disc ratios in early-type galaxies, 
going from parsec-sized nuclear discs, to kiloparsec-sized intermediate- and large-scale discs, 
in agreement with the findings of \citet{krajnovic2013}. 
Comparison of our results with those obtained by previous studies indicates that 
multicomponent models (as opposed to simple bulge/disc models) are necessary 
to derive reliable structural parameters, confirming previous results 
(e.g.~\citealt{laurikainen2005,laurikainen2007,laurikainen2010,lasker2014data,salo2015}). 
In general, the best-fit parameters obtained with 1D and 2D decomposition techniques for the same galaxy 
are consistent with each other, 
i.e.~no systematic effects were noticed between 1D and 2D modelling. 
However, our practical experience led us to prefer the 1D decomposition technique. 
Advantages associated with the 1D technique are a higher convergence rate for the fits, 
the wealth of information contained in the 1D isophotal analysis, 
and the easier interpretation of the 1D residuals. 
We caution against the dangerous practice of identifying unsubtracted galaxy components 
from the residual image of a 2D fit. 
As an additional warning, 
given the level of detail to which our galaxy decompositions were carried out, 
we do not consider possible for current automatic routines to reproduce our analysis. \\
future: bayes analysis to estimate max number of components, comparison with "physical" components; 
improving kin-phot approach, large kin maps are needed, 
combining hst + wide field obs to map galaxy outskirts, 
use bayes analysis to estimate uncertainties on params so that we don't need other decompositions, 
use of isofit and boggy's code for decomp, \\

@@

In Chapter \ref{ch:msigma}, we focused on the $M_{\rm BH} - \sigma$ diagram 
and in particular on the over-massive black holes that have been reported to exist 
at the high-mass end of the correlation. 
\citet{volontericiotti2013} explained the presence of the outlying black holes 
as a natural consequence of the fact that their host Central Cluster Galaxies (CCGs) have experienced more dry mergers 
than any other galaxy, 
since dry mergers are expected to increase the black hole mass 
and ``conserve'' the stellar velocity dispersion. 
We tested and disproved this scenario for 23 core-S\'ersic galaxies, 
using the ratio between black hole mass and stellar mass deficit as a proxy 
for the number of dry mergers experienced by a galaxy \citep{merritt2006}. 
This was confirmed by a second test using the central kinematic classification of 37 galaxies. 
Our analysis suggests that 
the merger history of the CCGs hosting over-massive black holes in the $M_{\rm BH} - \sigma$ diagram 
is not exceptional, 
therefore these galaxies should be considered as legitimate members of the $M_{\rm BH} - \sigma$ correlation. 
An alternative possibility for the CCGs is that their partially depleted core 
has been replenished by stars, globular clusters and nuclear star clusters 
acquired from minor dry mergers, 
where the accreted satellite galaxies do not host a massive black hole. 
However, high-resolution cosmological simulations seem to favour an ``inside-out'' formation scenario 
for early-type galaxies, 
where minor mergers only contribute to the build-up of the external envelope of high-mass galaxies 
(e.g.~\citealt{wellons2016}). 
It should also be noted that evidence is accumulating against the ``naive'' assumption 
that low-mass galaxies do not host any central massive black hole 
(e.g.~\citealt{Baldassare2015,Graham2016LEDA}). \\
shankar tbd

@@

(but see Shankar et al.~\emph{submitted to MNRAS} TBD). \\

>>>>

Chapter \ref{ch:galviv} \citep{paperI} was dedicated to the careful multicomponent decomposition of 66 galaxies 
with a direct measurement of the black hole mass. 
We followed the same approach as \citet{laurikainen2005} in selecting the components for each galaxy model.  
\emph{A priori} identification of the galaxy components was done on the basis of several different indicators 
such as the analysis of the isophotal parameters, 
the inspection of unsharp masks, 
complementary information extracted from the literature, 
and -- most importantly -- from the galaxy kinematics. 
A joint photometric-kinematic approach (e.g.~\citealt{krajnovic2013,arnold2014}) 
turned out to play a decisive role for the robustness of our galaxy modelling. 
Upon examining central \citep{atlas3dIII,scott2014} and more extended \citep{arnold2014} velocity maps 
and comparing kinematic and photometric signatures of stellar discs, 
we were able to securely identify the presence and the radial extent of such discs. 
We observed a wide range of spheroid-to-disc ratios in early-type galaxies, 
going from parsec-sized nuclear discs, to kiloparsec-sized intermediate- and large-scale discs, 
in agreement with the findings of \citet{krajnovic2013}. 
Comparison of our results with those obtained by previous studies indicates that 
multicomponent models (as opposed to simple bulge/disc models) are necessary 
to derive reliable structural parameters, confirming previous results 
(e.g.~\citealt{laurikainen2005,laurikainen2007,laurikainen2010,lasker2014data,salo2015}). 
In general, the best-fit parameters obtained with 1D and 2D decomposition techniques for the same galaxy 
are consistent with each other, 
i.e.~no systematic effects were noticed between 1D and 2D modelling. 
However, our practical experience led us to prefer the 1D decomposition technique. 
Advantages associated with the 1D technique are a higher convergence rate for the fits, 
the wealth of information contained in the 1D isophotal analysis, 
and the easier interpretation of the 1D residuals. 
We caution against the dangerous practice of identifying unsubtracted galaxy components 
from the residual image of a 2D fit. 
As an additional warning, 
given the level of detail to which our galaxy decompositions were carried out, 
we do not consider possible for current automatic routines to reproduce our analysis. \\
In our 1D galaxy decomposition code, the minimisation routine is based on the Levenberg-Marquardt algorithm, 
which is known to be one of the fastest algorithms available, 
but is also prone to ``get trapped'' into local minima of the chi-squared distribution. 
As a future project, it would be interesting to implement in the code different minimisation algorithms, 
such as those available in {\tt Imfit} \citep{imfit},  
and test their performance. 
It would also be beneficial to create a procedure to estimate the systematic uncertainties 
associated with the best-fit parameters from the fit itself, 
without the need to resort to decompositions performed by other authors as we did in this work. 
One of the issues associated with multicomponent decomposition is the risk of ``overfitting'', 
i.e.~having a model with more components (or parameters) than what is actually needed. 
Information theory and Bayes analysis provide us with a number of methods 
to estimate the maximum number of parameters required by a model 
to fit a given dataset, 
such as the \citet{akaike1974} Information Criterion that was used in Chapter \ref{ch:mn}. 
One could check whether or not these methods agree with our human-supervised selection 
of ``physically motivated'' model components. 
Our structural analysis of galaxies confirmed the usefulness of a joint photometric-kinematic approach, 
especially when extended (i.e.~$> 1 R_{\rm e}$, \citealt{arnold2014}) kinematic maps are available. 
Future galaxy decomposition studies will be able to take advantage from the large datasets 
of on-going Integral Field Spectroscopy surveys (e.g.~SAMI, \citealt{croom2012}, and MaNGA, \citealt{law2014}). 
Finally, we note that \citet{ciambur2015} recently developed the IRAF task {\tt isofit}, 
which allows the user to perform a more sophisticated isophotal analysis of galaxies than the task {\tt ellipse} 
in the presence of an inclined disc. 
Our isophotal analysis was carried out before {\tt isofit} was conceived or available, 
but future studies will be able to use it. \\

@@

In Chapter \ref{ch:msigma}, we focused on the $M_{\rm BH} - \sigma$ diagram 
and in particular on the over-massive black holes that have been reported to exist 
at the high-mass end of the correlation. 
\citet{volontericiotti2013} explained the presence of the outlying black holes 
as a natural consequence of the fact that their host Central Cluster Galaxies (CCGs) have experienced more dry mergers 
than any other galaxy, 
since dry mergers are expected to increase the black hole mass 
and ``conserve'' the stellar velocity dispersion. 
We tested and disproved this scenario for 23 core-S\'ersic galaxies, 
using the ratio between black hole mass and stellar mass deficit as a proxy 
for the number of dry mergers experienced by a galaxy \citep{merritt2006}. 
This was confirmed by a second test using the central kinematic classification of 37 galaxies. 
Our analysis suggests that 
the merger history of the CCGs hosting over-massive black holes in the $M_{\rm BH} - \sigma$ diagram 
is not exceptional, 
therefore these galaxies should be considered as legitimate members of the $M_{\rm BH} - \sigma$ correlation. 
An alternative possibility for the CCGs is that their partially depleted core 
has been replenished by stars, globular clusters and nuclear star clusters 
acquired from minor dry mergers, 
where the accreted satellite galaxies do not host a massive black hole. 
However, high-resolution cosmological simulations seem to favour an ``inside-out'' formation scenario 
for early-type galaxies, 
where minor mergers only contribute to the build-up of the external envelope of high-mass galaxies 
(e.g.~\citealt{wellons2016}). 
It should also be noted that evidence is accumulating against the ``naive'' assumption 
that low-mass galaxies do not host any central massive black hole 
(e.g.~\citealt{Baldassare2015,Graham2016LEDA}). \\

@@

%(but see Shankar et al.~\emph{submitted to MNRAS} TBD). \\

### 2016-03-10 thesis/mn.tex
conclusions ellicular
all chapters but concl

chapter x was dedicated to recovery of mn correlation with literature data

now we can do better with high quality data set

consistency with other correlations

good uncertainties -> robust estimate of intr scatter -> comparison with m-l

The remainder of this chapter comprises the published version of the paper 
``Supermassive Black Holes and Their Host Spheroids. 
III. The $M_{\rm BH} - n_{\rm sph}$ correlation'' 
by G.~A.~D.~Savorgnan,  
as it appears in Volume xxx of the \emph{The Astrophysical Journal}.

\includepdf[pages={1-10}]{mbhn.pdf}

>>>>

While the previous Chapter was dedicated to the study of the relations between 
black hole mass and galaxy or spheroid luminosity, 
in this Chapter we will perform an analogous analysis 
to explore the relation between black hole mass and spheroid S\'ersic index ($M_{\rm BH} - n_{\rm sph}$). 
To address the issue of consistency between galaxy scaling relations 
that was outlined in Chapter \ref{ch:intro}, 
it is mandatory to include also an analysis of the relation between spheroid luminosity 
and spheroid S\'ersic index ($L_{\rm sph} - n_{\rm sph}$). 
The reliability of the uncertainties associated with the S\'ersic index measurements 
obtained from our 1D decompositions 
ensures a robust estimate of the intrinsic scatter in the $M_{\rm BH} - n_{\rm sph}$ diagram, 
which can be compared with that in the $M_{\rm BH} - L_{\rm sph}$ diagram. \\

The remainder of this chapter comprises the peer-reviewed version of the paper 
``Supermassive Black Holes and Their Host Spheroids. 
III. The $M_{\rm BH} - n_{\rm sph}$ correlation'' 
by G.~A.~D.~Savorgnan,  
accepted for publication in the \emph{The Astrophysical Journal}.

\includepdf[pages={1-10}]{ApJ2016b.pdf}

### 2016-03-10 thesis/recovery-mn.tex
all chapters but concl

\cite{graham2001} demonstrated that the mass of SMBHs is correlated 
to the stellar light concentration of their host spheroid. 
Six years later, \cite{grahamdriver2007} expanded the galaxy sample of \cite{graham2001} 
and carried out 1D bulge/disc decompositions. 
They described the 1D surface brightness profile of spheroids with the S\'ersic model, 
and obtained a direct measure of the central radial concentration of stars 
by means of the S\'ersic index. 
\cite{grahamdriver2007} confirmed the early findings of \cite{graham2001}, 
presenting a strong correlation between the black hole mass and the spheroid S\'ersic index ($M_{\rm BH} - n_{\rm sph}$). 
They measured a small level of scatter, 
which made the $M_{\rm BH} - n_{\rm sph}$ and $M_{\rm BH} - \sigma_*$ correlations 
evenly competing for the title of fundamental black hole mass scaling relation. \\

As the number of directly measured black hole masses increased with time and
the constantly improving technology of machines allowed shorter computational times for 2D decomposition codes, 
more studies were dedicated to the analysis of various black hole mass scaling relations. 
\cite{sani2011}, \cite{vika2012}, and \cite{beifiori2012} attempted 2D decompositions 
of their samples of galaxies with a direct measure of the black hole mass. 
\citeauthor{sani2011} and \citeauthor{vika2012} included more than two components in their galaxy models 
-- such as bars and nuclear sources -- 
whereas \citeauthor{beifiori2012} excluded barred galaxies from their analysis. 
From these decompositions, they derived and analysed several correlations. 
However, none of these three studies was able to obtain a strong $M_{\rm BH} - n_{\rm sph}$ relation from their data. 
This raised an obvious question: 
what prevented \citeauthor{sani2011}, \citeauthor{vika2012}, and \citeauthor{beifiori2012} 
from the recovery of a tight $M_{\rm BH} - n_{\rm sph}$ correlation? 
Imputable factors could be the decomposition technique (1D versus 2D), 
the use of more sophisticated models featuring a larger number of components, 
the use of different wavelengths, 
or possibly the accuracy of the decompositions. \\

These circumstances motivated us to engage a preliminary study in 2012, 
aiming at explaining the lack of an $M_{\rm BH} - n_{\rm sph}$ correlation 
in the data of \citeauthor{sani2011}, \citeauthor{vika2012}, and \citeauthor{beifiori2012} 
After collecting and comparing the results from the galaxy decompositions 
of the aforementioned four studies (\citeauthor{grahamdriver2007}, \citeauthor{sani2011}, 
\citeauthor{vika2012}, and \citeauthor{beifiori2012}), 
we immediately noticed that some galaxies had been described with significantly different models 
(e.g.~the galaxy M60, treated as a discless elliptical galaxy by two studies, 
and as a lenticular galaxy by the other two studies). 
Different galaxy models (for the same galaxy) evidently resulted in different best-fit parameters. 
Not only this, but some studies obtained largely discrepant best-fit parameters 
for the same galaxy even though they used the same choice of decomposition 
(e.g.~two studies described the galaxy NGC 3115 with a S\'ersic-bulge + exponential-disc model, 
but obtained a bulge S\'ersic index of $3$ and $13$, respectively). 
This confirmed doubts about the accuracy of some decompositions. \\

For each galaxy, we averaged the available S\'ersic index measurements 
and used these average values to populate the $M_{\rm BH} - n_{\rm sph}$ diagram. 
Immediately, we noticed that a clear correlation was emerging from the data, 
although with a significant amount of scatter. 
We attributed the large amount of scatter to ``bad'' decompositions, 
and decided to develop a method for an impartial identification of the ``bad'' S\'ersic index measurements. 
After estimating a maximum tolerable disagreement between the S\'ersic index measurements of the same galaxy 
by taking into account the use of different decomposition methods and wavelengths, 
we excluded the most discrepant measurements and averaged the remaining ones. 
This resulted in a dramatic reduction of scatter 
and the recovery of a strong $M_{\rm BH} - n_{\rm sph}$ correlation. 
Finally, we explored substructure in the $M_{\rm BH} - n_{\rm sph}$ diagram 
expected for consistency with other known scaling relations. \\

>>>>

\cite{graham2001} demonstrated that the mass of SMBHs is correlated 
with the stellar light concentration of their host spheroid. 
Six years later, \cite{grahamdriver2007} expanded the galaxy sample of \cite{graham2001} 
and carried out 1D bulge/disc decompositions. 
They described the 1D surface brightness profile of spheroids with the S\'ersic model, 
and obtained a direct measure of the central radial concentration of stars 
by means of the S\'ersic index. 
\cite{grahamdriver2007} confirmed the early findings of \cite{graham2001}, 
presenting a strong correlation between the black hole mass and the spheroid S\'ersic index ($M_{\rm BH} - n_{\rm sph}$). 
They measured a small level of scatter, 
which made the $M_{\rm BH} - n_{\rm sph}$ and $M_{\rm BH} - \sigma_*$ correlations 
evenly competing for the title of fundamental black hole mass scaling relation. \\

As the number of directly measured black hole masses increased with time and
the constantly improving technology of machines allowed shorter computational times for 2D decomposition codes, 
more studies were dedicated to the analysis of various black hole mass scaling relations. 
\cite{sani2011}, \cite{vika2012}, and \cite{beifiori2012} attempted 2D decompositions 
of their samples of galaxies with a direct measure of the black hole mass. 
\citeauthor{sani2011} and \citeauthor{vika2012} included more than two components in their galaxy models 
such as bars and nuclear sources, 
whereas \citeauthor{beifiori2012} excluded barred galaxies from their analysis 
and used simple bulge+disc models. 
From these decompositions, they derived and analysed several correlations. 
However, none of these three studies was able to obtain a strong $M_{\rm BH} - n_{\rm sph}$ relation from their data. 
This raised an obvious question: 
what prevented \citeauthor{sani2011}, \citeauthor{vika2012}, and \citeauthor{beifiori2012} 
from the recovery of a tight $M_{\rm BH} - n_{\rm sph}$ correlation? 
Imputable factors could be the decomposition technique (1D versus 2D), 
the use of more sophisticated models featuring a larger number of components, 
the use of different wavelengths, 
or possibly the accuracy of the decompositions. \\

These circumstances motivated us to engage a preliminary study in 2012, 
aiming at explaining the lack of an $M_{\rm BH} - n_{\rm sph}$ correlation 
in the data of \citeauthor{sani2011}, \citeauthor{vika2012}, and \citeauthor{beifiori2012} 
After collecting and comparing the results from the galaxy decompositions 
of the aforementioned four studies (\citeauthor{grahamdriver2007}, \citeauthor{sani2011}, 
\citeauthor{vika2012}, and \citeauthor{beifiori2012}), 
we immediately noticed that some galaxies had been described by these authors with significantly different models 
(e.g.~the galaxy M60, treated as a discless elliptical galaxy by two studies, 
and as a lenticular galaxy by the other two studies). 
Different galaxy models (for the same galaxy) evidently resulted in different best-fit parameters. 
Not only this, but some studies obtained largely discrepant best-fit parameters 
for the same galaxy even when they used the same choice of decomposition 
(e.g.~two studies described the galaxy NGC 3115 with a S\'ersic-bulge + exponential-disc model, 
but obtained a bulge S\'ersic index of $3$ and $13$, respectively). 
This confirmed our doubts about the accuracy of some decompositions. \\

For each galaxy, we averaged the available S\'ersic index measurements 
and used these average values to populate the $M_{\rm BH} - n_{\rm sph}$ diagram. 
We noticed that a clear correlation was emerging from the data, 
although with a significant amount of scatter. 
We attributed the large amount of scatter to ``bad'' decompositions, 
and decided to develop a method for an impartial identification of the ``bad'' S\'ersic index measurements. 
After estimating a maximum tolerable disagreement between the S\'ersic index measurements of the same galaxy 
by taking into account the use of different decomposition methods and wavelengths, 
we excluded the most discrepant measurements and averaged the remaining ones. 
This resulted in a dramatic reduction of scatter 
and the recovery of a strong $M_{\rm BH} - n_{\rm sph}$ correlation. 
Finally, we explored substructure in the $M_{\rm BH} - n_{\rm sph}$ diagram 
expected for consistency with other known scaling relations. \\

### 2016-03-10 thesis/ubermassive-msigma.tex
conclusions ellicular
all chapters but concl

\chapter{m-sigma}
\label{ch:msigma}

A precise measurement of the log-slope of the $M_{\rm BH} - \sigma$ correlation is important 
to constrain theoretical models of AGN feedback. 
For example, energy driven outflows are expected to produce a scaling of $M_{\rm BH} \propto \sigma^5$, 
whereas momentum driven outflows should result into a shallower $M_{\rm BH} \propto \sigma^4$ relation 
\citep{silkrees1998,fabian1999}. 
Since the $M_{\rm BH} - \sigma$ correlation was presented for the first time by \citet{ferraresemerritt2000}
and \citet{gebhardt2000}, 
there has been an ongoing, lively debate about its log-slope, 
whose estimates differed by up to a few standard deviations 
depending on the choice of galaxy sample, the method used to measure the velocity dispersion,
the assumed uncertainty associated with the velocity dispersion, 
and the linear regression algorithm (e.g.~\citealt{tremaine2002}). \\

@@

$M_{\rm BH} - \sigma$ diagram \citep{mcconnell2011,mcconnell2012}, 

@@

Whether or not considering the over-massive black holes as an ``exception to the rule'', 
or in other words legitimately excluding them from the linear regression analysis of the $M_{\rm BH} - \sigma$ diagram, 
obviously has an impact on the estimate of the log-slope of the correlation. 
Therefore, it is important to test the scenario proposed by \citet{volontericiotti2013} 
with empirical data. \\

>>>>

\chapter{Monster black holes in Central Cluster Galaxies}
\label{ch:msigma}

A precise measurement of the log-slope of the $M_{\rm BH} - \sigma_*$ correlation is important 
to constrain theoretical models of AGN feedback. 
For example, energy driven outflows are expected to produce a scaling of $M_{\rm BH} \propto \sigma_*^5$, 
whereas momentum driven outflows should result into a shallower $M_{\rm BH} \propto \sigma_*^4$ relation 
\citep{silkrees1998,fabian1999}. 
Since the $M_{\rm BH} - \sigma_*$ correlation was presented for the first time by \citet{ferraresemerritt2000}
and \citet{gebhardt2000}, 
there has been an ongoing, lively debate about its log-slope, 
whose estimates differed by up to a few standard deviations 
depending on the choice of galaxy sample, the method used to measure the velocity dispersion,
the assumed uncertainty associated with the velocity dispersion, 
and the linear regression algorithm (e.g.~\citealt{tremaine2002}). \\

@@

$M_{\rm BH} - \sigma_*$ diagram \citep{mcconnell2011,mcconnell2012}, 

@@

Whether or not considering the over-massive black holes as an ``exception to the rule'', 
or in other words legitimately excluding them from the linear regression analysis of the $M_{\rm BH} - \sigma_*$ diagram, 
obviously has an impact on the estimate of the log-slope of the correlation. 
Therefore, it is important to test the scenario proposed by \citet{volontericiotti2013} 
with empirical data. \\
